---
---
---

# Scatter plot with the regression line can be added here for visualization.

### Apply Model to Test Data

# Evaluating the model using Mean Squared Error (MSE) on both training and test data

\\\\\\\<could interpret the Test MSE\\\\\\\>

```{r apply model to test_data}
# Calculating the training and test MSE for the linear regression model
train_mse = mean((train_data$medv - predict(lm.fit, train_data))^2)
test_mse = mean((test_data$medv - predict(lm.fit, test_data))^2)

print(paste("Training MSE:", round(train_mse, 2)))
print(paste("Test MSE:", round(test_mse, 2)))

```

### Simple Linear Regression results & interpretation

# Discussing how well the model fits the data and answering the question posed in the analysis

### Perform Multiple Linear Regression on Training Data

# Now adding 'age' as a predictor in the model to see if it improves prediction

\\\\\\\<what question does this model answer?\\\\\\\>

```{r}
# Fitting a multiple linear regression model to predict 'medv' from 'lstat' and 'age'
lm.multiple.fit = lm(medv ~ lstat + age, data = train_data)  
summary(lm.multiple.fit)
```

### Apply the Model to Test Data

```{r}
# Calculating training and test MSE for the multiple linear regression model
train_mse = mean((train_data$medv - predict(lm.multiple.fit, train_data))^2)
test_mse = mean((test_data$medv - predict(lm.fit, test_data))^2)

print(paste("Training MSE:", round(train_mse, 2)))
print(paste("Test MSE:", round(test_mse, 2)))



```

## NHANES Data Analysis

## Objective

# Describing the objective of predicting BMI for individuals between 18 and 70 using certain predictors

please predict BMI using Age, SmokeNow, PhysActive for individuals between the ages of 18 and 70

## Data understanding & Preparation

# Understanding the structure of the data to prepare it for analysis

## Data loading

```{r}
# Loading NHANES dataset and filtering for relevant columns and age range
#install.packages("NHANES")
library(NHANES)
data(NHANES)

SMOKERS = NHANES %>%
  select(BMI, Age, SmokeNow, PhysActive) %>%
  filter(Age >= 18 & Age <= 70)


```

## Data Exploration

```{r}
# Checking for missing values in the selected data
missing_values <- SMOKERS %>%
  summarise(across(everything(), ~ sum(is.na(.))))

print(missing_values)




```

## Train - Test Split

```{r}
# Setting a seed for reproducibility and splitting the SMOKERS data into training and test sets
# Add the 'id' column to the original dataset (SMOKERS)
SMOKERS <- SMOKERS %>% mutate(id = row_number())

# Split the data into 75% training and 25% testing
SMOKERS_split <- SMOKERS %>%
  sample_frac(0.75)  # 75% for training data

# The remaining 25% will be for testing
test_data <- anti_join(SMOKERS, SMOKERS_split, by = "id")  # Join using 'id' column
train_data <- SMOKERS_split



```

## Exploratory Data Analysis

```{r}
# Plotting the distribution of BMI values to understand its spread
ggplot(SMOKERS, aes(x = BMI)) +
  geom_histogram(fill = "steelblue", binwidth = 5, color = "white") +
  labs(title = "Distribution of BMI", x = "BMI", y = "Count")

```

## Scatterplot: Age vs. BMI

# Visualizing the relationship between Age and BMI

This helps in visualizing the relationship between **Age** and **BMI**.

```{r}
 #Scatterplot of Age vs. BMI
ggplot(SMOKERS, aes(x = Age, y = BMI)) +
  geom_point(alpha = 0.6, color = 'blue') +
  labs(title = "Scatterplot: Age vs. BMI", x = "Age", y = "BMI")


```

## Scatterplot: SmokeNow vs. BMI

```{r}
# Boxplot showing the relationship between Smoking status (SmokeNow) and BMI
ggplot(SMOKERS, aes(x = SmokeNow, y = BMI)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Boxplot: SmokeNow vs. BMI", x = "SmokeNow", y = "BMI")

```

## Model Implementation

### Perform simple Linear Regression on Training Data

```{r}
# Fitting a simple linear regression model to predict BMI from Age
#Fit a Simple Linear Regression Model ---
lm.fit <- lm(BMI ~ Age, data = train_data)

# View Model Summary
summary(lm.fit)

```

### Apply Model to Text Data

```{r}

# Calculating training and test MSE for the simple linear regression model
train_mse = mean((train_data$BMI - predict(lm.fit, train_data))^2)
test_mse = mean((test_data$BMI - predict(lm.fit, test_data))^2)

print(paste("Training MSE:", round(train_mse, 2)))
print(paste("Test MSE:", round(test_mse, 2)))

```

### Simple Linear Regression results & interpretation

# Interpretation of the results of the simple linear regression model

### Perform Multiple Linear Regression on Training Data

```{r}
# Fitting a multiple linear regression model to predict BMI using Age, SmokeNow, and PhysActive
lm.multiple.fit = lm(BMI ~ Age + SmokeNow + PhysActive, data = train_data)  
summary(lm.multiple.fit)


```

### Apply the Model to Test Data

```{r}
# Calculating the training and test MSE for the multiple linear regression model
train_mse = mean((train_data$BMI - predict(lm.multiple.fit, train_data))^2)
test_mse = mean((test_data$BMI - predict(lm.multiple.fit, test_data))^2)

print(paste("Training MSE:", round(train_mse, 2)))
print(paste("Test MSE:", round(test_mse, 2)))


```
